MS ANXIETY

### 2025-01-19 scripts/jama_scripts/MSAnxiety_manuscript_JAMA_revision_20250116.Rmd
w/steiger

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(stringr)
library(MatchIt)
library(ggpubr)
library(tidymv)
library(viridis)  # Ensure you have the viridis package
#functions
source("~/BBL/msanxiety/scripts/ms_functions.R")
```

@@

6. Supplement: 
 - Promis tables
#Rstduio Version: 2024.04.1+748 (2024.04.1+748)
## Make final data frame for analysis
### read in emr data
```{r read_in_df}
homedir <- "/Users/eballer/BBL/msanxiety/"

@@

#read in mimosa data
mimosa_qc <- read.csv("/Users/eballer/BBL/msdepression/data/melissa_martin_files/csv/mimosa_dataframe", sep = ",", header = FALSE) #n=2841

@@

fascicle_proportions <- read.csv(paste0("/Users/eballer/BBL/msdepression//results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
fascicle_proportions <- fascicle_proportions %>%
  mutate(EMPI = as.character(EMPI)) %>%
  mutate(EXAM_DATE = as.character(EXAM_DATE))

@@

lesion_volumes <- read.csv("/Users/eballer/BBL/msdepression/results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes", sep = ",", header = TRUE) #n=2336, 958 unique, identical to fascicle unique
lesion_volumes <- lesion_volumes %>%
  mutate(EMPI = as.character(EMPI)) %>%
  mutate(EXAM_DATE = as.character(EXAM_DATE))

@@

@@

#read in fiber volume values
fiber_volume_values <- read.csv("/Users/eballer/BBL/msdepression/templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv", sep=",", header = T)

@@

##Anxiety Network from neurosynth (association, not uniformity)
###read in file with fasciclexanxiety overlap, and define anxiety network
```{r read_in_neurosynth_anxiety_network}
neurosynth_anxiety_network_fascicle_overlap <- read.csv("/Users/eballer/BBL/msanxiety/results/streamline_volume_within_anxiety_network_association_neuro_maps_FDR_01.csv", header = T, sep = ",") %>% 
  filter(fascicle %in% fascicle_names_no_cranial_nerves) %>%
  slice(order(factor(fascicle, levels = fascicle_names_no_cranial_nerves))) %>% #get rid of cranial nerves up front 
  replace(is.na(.), 0)

@@

#this will display a demographics table comparing age, race, sex between healthy and depressed groups, using the first presentation to treatment
#msmeds simple, all groups
msmeds_table_all_groups <- make_ms_med_summary_table_ms_anxiety_simple(data_frame = df_demo_and_fascicles[df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified",])

#save 
msmeds_table_all_groups_df <- print(msmeds_table_all_groups, as.data.frame = TRUE)
write.csv(msmeds_table_all_groups_df, file = paste0(homedir, "/results/msmeds_table_all_groups_df.csv"), row.names = TRUE)

#msmeds simple, just MS+noA and MS+severeA
msmeds_table_diagnosis <- make_ms_med_summary_table_ms_anxiety_simple(data_frame = df_demo_and_fascicles[((df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") & (df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds")),])

#save 
msmeds_table_diagnosis_df <- print(msmeds_table_diagnosis, as.data.frame = TRUE)
write.csv(msmeds_table_diagnosis_df, file = paste0(homedir, "/results/msmeds_table_diagnosis_df.csv"), row.names = TRUE)
```

@@

```
####Test promis*anxiety dose  stats
#####Physical and Mental Health together
```{r phsy_and_mental_step_x_anxiety_dose}
###

@@

#read csv of empis from depression paper
depression_empis <- read.table(file="/Users/eballer/BBL/msanxiety/data/empis_of_people_in_BP_depression_paper.csv")
names(depression_empis) <- "EMPI"

@@

@@

#### Sensitivity
###### UF by depGroupVar (vol and prop, N.S)
```{r UF_by_depGroupVar}
# Specificity - want to make sure the UF findings are specific to anxiety, so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

@@

  #print summary stats
 gam_summary_stats(dep_mean_UF_gam_prop)
```

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(visreg)
library(lme4)
library(lmerTest)
library(ggplot2)
library(reshape2)
library(pwr)
library(mgcv)
library(stringr)
library(MatchIt)
library(ggpubr)
library(tidygam)
library(psych)
#library(tidymv)
library(aod)
library(viridis)  # Ensure you have the viridis package
#functions
source("~/Library/CloudStorage/Box-Box/BBL/ms_generic_scripts/msdepression_functions.R")
```

@@

6. Supplement: 
 - Promis tables
#Rstduio Version: 2024.04.1+748 (2024.04.1+748)
## Make final data frame for analysis
### read in emr data
```{r read_in_df}
homedir <- "~/Library/CloudStorage/Box-Box/BBL/msanxiety/"

@@

#read in mimosa data
mimosa_qc <- read.csv("~/Library/CloudStorage/Box-Box/BBL/msdepression/data/melissa_martin_files/csv/mimosa_dataframe", sep = ",", header = FALSE) #n=2841

@@

fascicle_proportions <- read.csv(paste0("~/Library/CloudStorage/Box-Box/BBL/msdepression//results/fascicle_volumes_all_subjects_roi_n2336.csv"), header = T, sep = ",") #n = 2336
fascicle_proportions <- fascicle_proportions %>%
  mutate(EMPI = as.character(EMPI)) %>%
  mutate(EXAM_DATE = as.character(EXAM_DATE))

@@

lesion_volumes <- read.csv("~/Library/CloudStorage/Box-Box/BBL/msdepression/results/mimosa_binary_masks_hcp_space_20211026_n2336_volumes", sep = ",", header = TRUE) #n=2336, 958 unique, identical to fascicle unique
lesion_volumes <- lesion_volumes %>%
  mutate(EMPI = as.character(EMPI)) %>%
  mutate(EXAM_DATE = as.character(EXAM_DATE))

@@

### read in brain volume size
```{r read_in_brain_volumes}
##########################################################
#             Read in Brain Volumes                     #
##########################################################

brain_volumes <- read.csv(paste0(homedir, "/results/total_fast_brain_volumes_all_subjs.csv"), sep = ",", header = TRUE) #n=2336, 
brain_volumes <- brain_volumes %>%
  mutate(EMPI = as.character(EMPI)) %>%
  mutate(EXAM_DATE = as.character(EXAM_DATE))

### merge brain vol with data frame with good mimosas
```{r merge_brain_volume_w_mimosa_df}
##########################################################
#             Merge lesion volumes with mimosa df        #
##########################################################
#merge whole data set with fascicle proportions, n = 814
df_demo_and_fascicles <- df_demo_and_fascicles %>%
  left_join(brain_volumes, by = c("EMPI", "EXAM_DATE"))

#test to make sure age is correlated with gm and tbv
gmv_cor_age <- cor.test(df_demo_and_fascicles$PAT_AGE_AT_EXAM, df_demo_and_fascicles$gm_volume)
print(gmv_cor_age) #Yup: t = -6.9769, df = 812, p-value = 6.264e-12 (higher age, lower gmv); cor = -0.237817

tbv_cor_age <- cor.test(df_demo_and_fascicles$PAT_AGE_AT_EXAM, df_demo_and_fascicles$total_volume)
print(tbv_cor_age) #Yup: t = -4.5874, df = 812, p-value = 5.196e-06 (higher age, lower tbv); cor = -0.158941 
```

@@

#read in fiber volume values
fiber_volume_values <- read.csv("~/Library/CloudStorage/Box-Box//BBL/msdepression/templates/dti/HCP_YA1065_tractography/fiber_volume_values.csv", sep=",", header = T)

@@

##Anxiety Network from neurosynth (association, not uniformity)
###read in file with fasciclexanxiety overlap, and define anxiety network
```{r read_in_neurosynth_anxiety_network}
neurosynth_anxiety_network_fascicle_overlap <- read.csv(paste0(homedir, "/results/streamline_volume_within_anxiety_network_association_neuro_maps_FDR_01.csv"), header = T, sep = ",") %>% 
  filter(fascicle %in% fascicle_names_no_cranial_nerves) %>%
  slice(order(factor(fascicle, levels = fascicle_names_no_cranial_nerves))) %>% #get rid of cranial nerves up front 
  replace(is.na(.), 0)

@@

#this will display a demographics table comparing age, race, sex between healthy and depressed groups, using the first presentation to treatment
#msmeds simple, all groups
#msmeds_table_all_groups <- make_ms_med_summary_table_ms_anxiety_simple(data_frame = df_demo_and_fascicles[df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified",])

#save 
#msmeds_table_all_groups_df <- print(msmeds_table_all_groups, as.data.frame = TRUE)
#write.csv(msmeds_table_all_groups_df, file = paste0(homedir, "/results/msmeds_table_all_groups_df.csv"), row.names = TRUE)

#msmeds simple, just MS+noA and MS+severeA
#msmeds_table_diagnosis <- make_ms_med_summary_table_ms_anxiety_simple(data_frame = df_demo_and_fascicles[((df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") & (df_demo_and_fascicles$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds")),])

#save 
#msmeds_table_diagnosis_df <- print(msmeds_table_diagnosis, as.data.frame = TRUE)
#write.csv(msmeds_table_diagnosis_df, file = paste0(homedir, "/results/msmeds_table_diagnosis_df.csv"), row.names = TRUE)
```

@@

```
####Test promis*anxiety dose  stats
#####NEW JAMA ANALYSIS: Testing for Normality - all normally distributed
```{r normality_testing}
###

### Normality tests

### MS noA
## MS+noA, Physical -> W = 0.87548, p-value = 0.07673 (normal)
ms_noa_physical_normality <- shapiro.test(df_for_promis_plot_mental_vs_physical_healthy_test$physical_health_score[df_for_promis_plot_mental_vs_physical_healthy_test$anxiety_dose == 0])
print(ms_noa_physical_normality)
## MS+noA, Emotional -> W = 0.92006, p-value = 0.2864 (normal)
ms_noa_emotional_normality <- shapiro.test(df_for_promis_plot_mental_vs_physical_healthy_test$mental_health_score[df_for_promis_plot_mental_vs_physical_healthy_test$anxiety_dose == 0])
print(ms_noa_emotional_normality)

### MS MILD
## MS+mildA, Physical -> W = 0.96068, p-value = 0.1297 (normal)
ms_milda_physical_normality <- shapiro.test(df_for_promis_plot_mental_vs_physical_healthy_test$physical_health_score[df_for_promis_plot_mental_vs_physical_healthy_test$anxiety_dose == 1])
print(ms_milda_physical_normality)
## MS+mildA, Emotional -> W = 0.97038, p-value = 0.299 (normal)
ms_milda_emotional_normality <- shapiro.test(df_for_promis_plot_mental_vs_physical_healthy_test$mental_health_score[df_for_promis_plot_mental_vs_physical_healthy_test$anxiety_dose == 1])
print(ms_milda_emotional_normality)

### MS SEVERE
## MS+severeA, Physical -> W = 0.96162, p-value = 0.8325 (normal)
ms_severea_physical_normality <- shapiro.test(df_for_promis_plot_mental_vs_physical_healthy_test$physical_health_score[df_for_promis_plot_mental_vs_physical_healthy_test$anxiety_dose == 2])
print(ms_severea_physical_normality)
## MS+severeA, Emotional -> W = 0.89898, p-value = 0.3248 (normal)
ms_severea_emotional_normality <- shapiro.test(df_for_promis_plot_mental_vs_physical_healthy_test$mental_health_score[df_for_promis_plot_mental_vs_physical_healthy_test$anxiety_dose == 2])
print(ms_severea_emotional_normality)

normality_stats_table <- data.frame(matrix(nrow = 6, ncol = 5))
names(normality_stats_table) <- c("Anxiety Severity", "PROMIS Functioning Domain", "W", "P", "Normality")
normality_stats_table$`Anxiety Severity` <- rep(c("MS+noA", "MS+mildA", "MS+severeA"), 2)
normality_stats_table$`PROMIS Functioning Domain` <- rep(c("Physical", "Emotional"), 3)

count = 1
for (anxiety_dose in 0:2){
  for (promis in c("physical_health_score", "mental_health_score"))
  {
    normality_test <- eval(parse(text = paste0("shapiro.test(df_for_promis_plot_mental_vs_physical_healthy_test$", promis, "[df_for_promis_plot_mental_vs_physical_healthy_test$anxiety_dose == ", anxiety_dose, "])")))
    print(normality_test)
    normality_stats_table$W[count] = round(normality_test$statistic, 3)
    normality_stats_table$P[count] = round(normality_test$p.value, 3)
    normality_stats_table$Normality = ifelse(normality_test$p.value < 0.05, "NON-NORMAL", "NORMAL")
    count = count + 1
  }
}
print(normality_stats_table)

write.csv(normality_stats_table, paste0(homedir, "results/normality_table_promis.csv"), row.names = F, sep = ",", col.names = T)

#####Physical and Mental Health together
```{r phsy_and_mental_step_x_anxiety_dose}
###

@@

#read csv of empis from depression paper
depression_empis <- read.table(file=paste0(homedir, "/data/empis_of_people_in_BP_depression_paper.csv"))
names(depression_empis) <- "EMPI"

@@

### Whole brain volumes 
#### Diagnosis 
  - sick > healthy in vol of lesions
```{r lesion_volumes_dx}
 df_empi_dx_network_proportions <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, total_volume) %>%
  ungroup() #n=123

#gam brain volume; DiagnosisMS+Anxiety     3359      25590   0.131    0.896    
#(T = 0.131, P = 0.896, Cohen's f2 =  0.4, Effect size = large)
total_brain_volume_gam <- gam(total_volume ~ Diagnosis + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data = df_empi_dx_network_proportions)
summary(total_brain_volume_gam)

 #print summary stats
 gam_summary_stats(total_brain_volume_gam)
```

#### Anxiety dose
  - anxiety dose correlates with disease burden (total brain volume )
```{r lesion_volumes_anxiety_dose}
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter (anxiety_dose != -1) #n=372

 #brain volume anxiety dose + sex + age: anxiety_dose     3948      11484   0.344    0.731  
#(T = 0.344, P = 0.731, Cohen's f2 =  0.311, Effect size = medium)

  anxiety_dose_total_brain_volume_by_sex_gam <- gam(total_volume ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = T), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxiety_dose_total_brain_volume_by_sex_gam)

   #print summary stats
 gam_summary_stats(anxiety_dose_total_brain_volume_by_sex_gam)

@@

#### Sensitivity
###### UF by depGroupVar (vol/prop, N.S)
```{r UF_by_depGroupVar}
# Specificity - want to make sure the UF findings are specific to anxiety, so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

@@

  #print summary stats
 gam_summary_stats(dep_mean_UF_gam_prop)

###### Lesion & brainvol by depGroupVar (lesion vol - Y, tbv - N, N.S)
```{r UF_by_depGroupVar}
# Specificity - want to make sure the lesion and brain vol findings are nonspecific t so check to see if same findings exist in depression
### Reviewed previous pull from BP paper, finding was also consistent there -> no diff in uncinate in depressed vs non depressed

 #mean vol mimosa lesions, depDiagnosisMS+noDepression 
#(T = -3.164, P = 0.002, Cohen's f2 =  0.128, Effect size = small)
dep_mean_lesion_vol_gam_prop <- gam(volume_of_mimosa_lesions ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_lesion_vol_gam_prop)

  #print summary stats
 gam_summary_stats(dep_mean_lesion_vol_gam_prop)

  #mean brain vol mimosa lesions, depDiagnosisMS+noDepression  
 #(T = 0.016, P = 0.988, Cohen's f2 =  0.333, Effect size = medium)
dep_mean_total_brain_vol_gam_prop <- gam(total_volume ~ depDiagnosis + osex + s(PAT_AGE_AT_EXAM, k=4, fx=F), data = df_dep)
summary(dep_mean_total_brain_vol_gam_prop)

  #print summary stats
 gam_summary_stats(dep_mean_total_brain_vol_gam_prop)

@@

###### JAMA ANALYSIS- UF Anxiety diagnosis + lesion vol
```{r uncinate_by_dx}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

### Results:  
#define group
  uncinate_by_dx_df <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(mean_UF_vol_scaled_total_vol_lesions = mean_UF_vol/volume_of_mimosa_lesions) %>%
  ungroup() #n=123

#bilateral UF ~ Diagnosis: DiagnosisMS+Anxiety   0.10624    0.05255   2.022   0.0455  * original result

#First, test if they are related. If so, they can't really be modeled together ->

  corr_LV_to_UF <- cor.test(uncinate_by_dx_df$volume_of_mimosa_lesions,uncinate_by_dx_df$mean_UF_vol) #corr 0.69; 
  print(corr_LV_to_UF)

  lesion_vol_UF_lm <- lm(uncinate_by_dx_df$volume_of_mimosa_lesions~ uncinate_by_dx_df$mean_UF_vol) #P <2.2*10^-16, they are colinear
  summary(lesion_vol_UF_lm) #T=10.585

  #how about in fornix

  corr_LV_to_fornix <- cor.test(uncinate_by_dx_df$volume_of_mimosa_lesions,uncinate_by_dx_df$mean_fornix_volume) #corr 0.69; 
  print(corr_LV_to_fornix) #0.3452654

  lesion_vol_fornix_lm <- lm(uncinate_by_dx_df$volume_of_mimosa_lesions~ uncinate_by_dx_df$mean_fornix_volume)
  summary(lesion_vol_fornix_lm) #T = 4.047

  ## Try scaling mean volume
  disease_mean_scaled_UF_vol_gam <- gam(mean_UF_vol_scaled_total_vol_lesions ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
  summary(disease_mean_scaled_UF_vol_gam) #- N.S.

  #print summary stats; (T = -0.556, P = 0.579, Cohen's f2 =  0.007, Effect size = minimal)"
  gam_summary_stats(disease_mean_scaled_UF_vol_gam)

  #How about predicting Diagnosis? - nope
  diagnosis_prediction <- glm(Diagnosis ~ mean_UF_prop + Diagnosis + osex+PAT_AGE_AT_EXAM + volume_of_mimosa_lesions, data = uncinate_by_dx_df, family = "binomial")
  summary(diagnosis_prediction)
  confint(diagnosis_prediction)

  #how about predicting diagnosis with scaled UF? N.S.
  diagnosis_prediction_scaled <- glm(Diagnosis ~ mean_UF_vol_scaled_total_vol_lesions + Diagnosis + osex+PAT_AGE_AT_EXAM, data = uncinate_by_dx_df, family = "binomial")
  summary(diagnosis_prediction_scaled)
  confint(diagnosis_prediction_scaled)

## modeling this doesn't make sense
#disease_mean_UF_prop_w_LV_gam <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F) + volume_of_mimosa_lesions, data = uncinate_by_dx_df)
#summary(disease_mean_UF_prop_w_LV_gam)

 #print summary stats
 # gam_summary_stats(disease_mean_UF_prop_w_LV_gam)
```
###### JAMA ANALYSIS-  UF Anxiety dose + lesion vol
```{r uncinateByAnxietyDose}

    #bilat anxiety dose + sex + age: anxiety_dose    0.04989    0.02397   2.082   0.0381 * original result

  #How about predicting Diagnosis? volume_of_mimosa_lesions  1.518e-05  7.147e-06   2.124   0.0357 *
  diagnosis_prediction <- lm(anxiety_dose ~ mean_UF_prop + osex+PAT_AGE_AT_EXAM + volume_of_mimosa_lesions, data = uncinate_by_dx_df)
  summary(diagnosis_prediction)
  confint(diagnosis_prediction)

### modeling volume of mimosa lesions like this doesn't make sense
 #anxietyGrouping_mean_UF_by_sex_w_LV_gam_prop <- gam(mean_UF_prop ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F) + volume_of_mimosa_lesions, data=df_demo_and_fascicles_no_unclass_anxiety_dose)
 # summary(anxietyGrouping_mean_UF_by_sex_w_LV_gam_prop)

  #print summary stats
  #gam_summary_stats(anxietyGrouping_mean_UF_by_sex_w_LV_gam_prop)

```
###### JAMA ANALYSIS- UF Anxiety diagnosis + total brain vol
```{r uncinate_by_dx}
### Having found that there is no association between the number of uncinate lesions and the anxiety dose, I next explored whether the *burden* of uncinate lesions differs by diagnosis/group

#bilateral UF ~ Diagnosis: DiagnosisMS+Anxiety   0.10624    0.05255   2.022   0.0455  * original result

#First, test if they are related. If so, they can't really be modeled together ->

 # t = -0.34681, df = 121, p-value = 0.7293; corr -0.03151255

  corr_tbv_to_UF <- cor.test(uncinate_by_dx_df$total_volume,uncinate_by_dx_df$mean_UF_vol) #corr 0.69; 
  print(corr_tbv_to_UF)

  total_vol_UF_lm <- lm(uncinate_by_dx_df$total_volume~ uncinate_by_dx_df$mean_UF_vol) #P <2.2*10^-16, they are colinear
  summary(total_vol_UF_lm)

  #can covary, no change: "(T = 2.009, P = 0.047, Cohen's f2 =  0.19, Effect size = medium)"
  disease_UF_cov_brain_vol_gam <- gam(mean_UF_prop ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F) + total_volume, data = uncinate_by_dx_df)
  summary(disease_UF_cov_brain_vol_gam) #- N.S.
  gam_summary_stats(disease_UF_cov_brain_vol_gam)
```

###### JAMA ANALYSIS-  UF Anxiety dose + total brain volume
```{r uncinateByAnxietyDose}

#check correlations  
# cor: -0.06590961 ; t = -1.2706, df = 370, p-value = 0.2047
  corr_tbv_to_UF <- cor.test(df_demo_and_fascicles_no_unclass_anxiety_dose$total_volume,df_demo_and_fascicles_no_unclass_anxiety_dose$mean_UF_vol) #corr 0.69; 
  print(corr_tbv_to_UF)

  total_vol_UF_lm <- lm(df_demo_and_fascicles_no_unclass_anxiety_dose$total_volume~ df_demo_and_fascicles_no_unclass_anxiety_dose$mean_UF_vol) #P <2.2*10^-16, they are colinear
  summary(total_vol_UF_lm)

  #can covary, no change: (T = 2.089, P = 0.037, Cohen's f2 =  0.102, Effect size = small)
  disease_UF_cov_brain_vol_dose_gam <- gam(mean_UF_prop ~ anxiety_dose + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F) + total_volume, data = df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(disease_UF_cov_brain_vol_dose_gam) #- N.S.
  gam_summary_stats(disease_UF_cov_brain_vol_dose_gam)

```
##### JAMA Analysis: Residuals vs UF
```{r dep_anxiety_UF}

## Definte data frams
################
## depression ##
################
df_dep <- df_demo_and_fascicles %>%
  filter(depGroupVar != 0) %>%
  mutate(depDiagnosis = as.factor(ifelse(depGroupVar ==2, "MS+Depression", "MS+noDepression")))

####################### 
## anxiety diagnosis ##
#######################
df_anxiety_dx <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(anxiety_Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(anxiety_Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(anxiety_Diagnosis = as.factor(anxiety_Diagnosis)) %>%
  ungroup()

##################
## anxiety dose ##
################## 
df_anxiety_dose <-  df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) #n=372

### stats
# Step 1: Fit the three models

# Model 1: depDiagnosis with df_dep
model_depDiagnosis <- gam(mean_UF_prop ~ depDiagnosis + s(PAT_AGE_AT_EXAM, k=4, fx=F) + osex, data = df_dep)

# Model 2: anxiety_Diagnosis with df_anxiety_dx
model_anxiety_Diagnosis <- gam(mean_UF_prop ~ anxiety_Diagnosis + s(PAT_AGE_AT_EXAM, k=4, fx=F) + osex, data = df_anxiety_dx)

# Model 3: anxiety_dose with df_anxiety_dose
model_anxiety_dose <- gam(mean_UF_prop ~ anxiety_dose + s(PAT_AGE_AT_EXAM, k=4, fx=F) + osex, data = df_anxiety_dose)

# Step 2: Extract residuals from each model (mean_UF_prop adjusted for age and sex)
residuals_depDiagnosis <- residuals(model_depDiagnosis)
residuals_anxiety_Diagnosis <- residuals(model_anxiety_Diagnosis)
residuals_anxiety_dose <- residuals(model_anxiety_dose)

# Step 3: Calculate the correlation between mean_UF_prop and the predictors (residuals)
cor_depDiagnosis <- cor(df_dep$mean_UF_prop, residuals_depDiagnosis)
cor_anxiety_Diagnosis <- cor(df_anxiety_dx$mean_UF_prop, residuals_anxiety_Diagnosis)
cor_anxiety_dose <- cor(df_anxiety_dose$mean_UF_prop, residuals_anxiety_dose)

# Print the correlation results
cat("Correlation between mean_UF_prop and depDiagnosis: ", cor_depDiagnosis, "\n")
cat("Correlation between mean_UF_prop and anxiety_Diagnosis: ", cor_anxiety_Diagnosis, "\n")
cat("Correlation between mean_UF_prop and anxiety_dose: ", cor_anxiety_dose, "\n")

# Step 4: Perform Steiger's test for comparing the correlations
# Create a function for Steiger's test (comparison of two correlated correlations)
steiger_test <- function(r12, r13, n) {
  # r12: correlation between variable1 and variable2
  # r13: correlation between variable1 and variable3
  # n: sample size

  # Calculate the Z score for comparison
  z <- (r12 - r13) / sqrt((1 - r12^2) / (n - 3) + (1 - r13^2) / (n - 3))
  p_value <- 2 * (1 - pnorm(abs(z)))  # Two-tailed p-value

  return(list(Z_score = z, p_value = p_value))
}

# Sample size (number of subjects)
n = length(Reduce(intersect, list(df_dep$EMPI, df_anxiety_dx$EMPI, df_anxiety_dose$EMPI)))

# Step 5: Test for differences between correlations
# Test if the correlation between anxiety_Diagnosis and mean_UF_prop is stronger than depression
steiger_test_anxiety_vs_dep <- steiger_test(cor_anxiety_Diagnosis, cor_depDiagnosis, n)

# Test if the correlation between anxiety_dose and mean_UF_prop is stronger than depression
steiger_test_anxiety_dose_vs_dep <- steiger_test(cor_anxiety_dose, cor_depDiagnosis, n)

# Output the results of Steiger's test; Z score:  -0.7806441 ; p-value:  0.4350118 
cat("Steiger's Test: Anxiety Diagnosis vs Depression Diagnosis\n")
cat("Z score: ", steiger_test_anxiety_vs_dep$Z_score, "\n")
cat("p-value: ", steiger_test_anxiety_vs_dep$p_value, "\n")

#Z score:  0.07445476 ; p-value:  0.9406485 
cat("Steiger's Test: Anxiety Dose vs Depression Diagnosis\n")
cat("Z score: ", steiger_test_anxiety_dose_vs_dep$Z_score, "\n")
cat("p-value: ", steiger_test_anxiety_dose_vs_dep$p_value, "\n")
```

```
##### JAMA ANALYSIS - compare correlations w/wout uncinate
```{r other_injured_fascicles}
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1)  %>% #n=372 
  mutate(total_lesion_burden_no_UF = sum_fascicle_vol_lost - (UF_L_vol + UF_R_vol))

#almost identical because uncinate is small
#t=2.05, p = 0.04083
cor_lesion_burden_by_anxiety = cor.test(df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose, df_demo_and_fascicles_no_unclass_anxiety_dose$sum_fascicle_vol_lost)
print(cor_lesion_burden_by_anxiety)

#t=2.05, p = 0.04075
cor_lesion_burden_no_UF_by_anxiety = cor.test(df_demo_and_fascicles_no_unclass_anxiety_dose$anxiety_dose, df_demo_and_fascicles_no_unclass_anxiety_dose$total_lesion_burden_no_UF)
print(cor_lesion_burden_no_UF_by_anxiety)

@@

```{r ants}
